---
title: "Testing doses and supplements in R tooth growth data"
author: "Marc T. Henry de Frahan"
output: 
  pdf_document
---

# Synopsis

# Load some libraries
```{r}
library(dplyr)
library(ggplot2)
```
# Exploratory Data Analysis

Let's first load the data and get some basic information from the data
set. We also transform the dose variable to a factor, which will be
easier to use later on.

```{r loading}
df <- ToothGrowth
df <- df %>%
    mutate(dose = as.factor(dose)) %>%
	rename(delivery = supp)
str(df)
```

We are looking at `r nrow(df)` observations of tooth length where
there are `r nlevels(df$delivery)` delivery methods (orange juice (OJ) and
ascorbic acid (VC)) and `r nlevels(df$dose)` dose levels (`r
levels(df$dose)`).

## Visualisation of the effect of the dose amount on tooth growth

As shown in this figure, there might be an effect of the dose amount
on the tooth growth. We will explore this in more detail later.

```{r boxplot_dose}
ggplot(data=df, aes(x=dose,y=len)) +
    geom_boxplot(aes(fill=dose)) +
    labs(x = "dose",
         y="Tooth length",
         title="Boxplot of tooth length versus dose");
```
## Visualisation of the effect of the delivery method on tooth growth

As shown in this figure, there does not seem to be much of an effect
of the delivery method on tooth growth. We will explore this in more
detail later.

```{r boxplot_delivery}
ggplot(data=df, aes(x=delivery,y=len)) +
    geom_boxplot(aes(fill=delivery)) +
    labs(x = "delivery",
         y="Tooth length",
         title="Boxplot of tooth length versus delivery method");
```


# Questions and Key Assumptions

Our analysis will seek to answer two questions:

1. Does the dose amount change the tooth growth?
2. Does the delivery method change the tooth growth?

For our analysis to hold, we make the following key assumptions about
the data:

- 

# Hypothesis Tests 

# Conclusions 



\newpage

# Appendix

<!-- ```{r code=readLines(knitr::purl('./tooth_growth.Rmd', documentation = 1)), eval = FALSE} -->
<!-- ``` -->
